---
title: "Scoring fly enhancers using NRLB models"
author: "Chaitanya Rastogi and Harmen J. Bussemaker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scoring fly enhancers using NRLB models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette shows how to use the NRLBtools package. It is related to the paper on the No Read Left Behind (NRLB) algorithm by Rastogi et al., PNAS (2018).

## Working with NRLB models for DNA binding specificity
The \code{NRLBtools} package makes it easy to use the DNA binding specicity models that were built from SELEX in vitro binding data using the NRLB algorithm and as described in the Rastogi et al., PNAS (2018) paper. The available models are:
```{r}
library(NRLBtools)
NRLBtools::model.info()
```
## Energy Logo
It is always a good idea to visualize the model as an energy logo before using it. We can visualize the Exd-UbxIVa model used in the paper by using the logo function and its model number (16 in this case):
```{r}
NRLBtools::logo(nrlb.model = 16)
```
We can actually use model.info to provide more information about the NRLB models that come with this package. For example, we can see that model number 16 is a multi-mode model with two modes:
```{r}
NRLBtools::model.info(nrlb.model = 16)
```

## Visualizing affinity profiles for enhancers
We start by extracting the sequence underlying the genomic region of interest. Here, we will only use the E3N enhancer element (as shown in the paper) from the dm3 Drosophila genome: 
```{r}
enh = as.character(BSgenome.Dmelanogaster.UCSC.dm3::BSgenome.Dmelanogaster.UCSC.dm3$chrX[4915195:4915486])
```
An affinity profile along the E3N enhancer can be created as follows:
```{r, fig.show='asis'}
NRLBtools::score.seq(sequence = enh, nrlb.model = 16, plot = TRUE, nPeaks = 10, annotate = TRUE)
```

## Working with new models
The NRLBtools package can also be used to interpret new NRLB models fit to SELEX data. To do so, one first needs to download NRLB from https://github.com/BussemakerLab/NRLB. For the following example, we will be using the precomputed results of running the default fit detailed in the README file. We begin by loading the CSV model results file, which contains all the relevant model information in a machine-readable format:
```{r}
max = NRLBtools::load.models(fileName = system.file("extdata", "MAX-NRLBConfig.csv", package = "NRLBtools"))
```
Next, we can use model info to explore the types of models that are stored in this object:
```{r}
NRLBtools::model.info(models = max)
```
We can get more information about a particular model by providing the model's row index:
```{r}
NRLBtools::model.info(models = max, index = 8)
```
We can also visualize our new model:
```{r}
NRLBtools::logo(models = max, index = 8)
```
And use the model to score a sequence:
```{r}
NRLBtools::score.seq(sequence = enh, models = max, index = 8, plot = T)
```